---
title: "Data607 - Global Baseline Estimates"
author: "Anthony Josue Roman"
date: "2024-09-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
require(readr)
require(dplyr)
require(tidyr)
require(tidyverse)
require(ggplot2)
require(knitr)
require(kableExtra)

library(readr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
```

This assignment will use the following packages:

```{r eval=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
```
## General Overview

This assignment for extra credit will utilize the data from Assignment 2 from my [GitHub page](https://github.com/spacerome/Data607_Assignment_2).

This assignment will also be available on my [GitHub Repo](https://github.com/spacerome/Data607_EC_GBE).

The purpose of this assignment is to utilize Global Baseline estimates to predict what critics would rate films that they did not see. Similar to what I mentioned on my previous assignment, there will be a scale of 1 to 5, where 1 is the lowest and 5 is the highest. If and only if there are critics that did not see such movie, the data will output as `NA`. The data set utilizes a random algorithm, which has each of the random 20 names generated to "rate" the 20 videos picked from the dataset taken from [MovieLens](https://grouplens.org/datasets/movielens/) and output a ratings file. The file was then combined into `movieratings.csv` for this assignment for simplification purposes.

### Movie Ratings File Extraction

We will first extract the data from the `movie_ratings.csv` file to gather the required variables from Assignment 2 utilizing the following code block:

```{r echo =TRUE}

# Initialize directory
a <- getwd()
setwd(a)

movieRatings <- read.csv("movie_ratings.csv")

```

### Initializing Global Mean

Here we have the `globalMean` equate to `2.965`. 

```{r echo =TRUE}

globalMean <- mean(movieRatings$rating, na.rm = TRUE)
print(globalMean)

```

### Critics vs Global Mean

The following table uses the following values to show the average ratings of each user, and the User Average subtracted by the `globalMean`. The `globalMeanUserAvg` subtracts `userAvg` with `globalMean`. This will be used further in the assignment. The number of digits was condensed into a format of `#.##` to make the tables a bit smaller.

```{r echo=TRUE}

averageUserRating <- movieRatings %>%
  group_by(userName) %>%
  summarise(userAvg = mean(rating, na.rm = TRUE)) %>%
  mutate(globalMeanUserAvg = userAvg - globalMean)

averageUserRating %>%
  kable(col.names = c("Critic", "Average Rating", "User Average - Mean Movie"), digits = 2, align = 'lcc') %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

```

### Movies vs Global Mean

The following table uses the following values to show the average ratings of each movie, and the Movie Average subtracted by the `globalMean`. The `globalMeanMovieAvg` subtracts `MovieAvg` with `globalMean`. This will be used further in the assignment. The number of digits was condensed into a format of `#.##` to make the tables a bit smaller.

```{r echo=TRUE}

averageUserRating <- movieRatings %>%
  group_by(movieName) %>%
  summarise(movieAvg = mean(rating, na.rm = TRUE)) %>%
  mutate(globalMeanMovieAvg = movieAvg - globalMean)

averageUserRating %>%
  kable(col.names = c("Movie", "Average Rating", "Movie Average - Mean Movie"), digits = 2, align = 'lcc') %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

```

### Generating the Tables

This code block will prepare the table to for the global base line estimate. Since the movie title numbers is greater than 10, I decided to split it apart to make it better visually. I also included a section to calculate the `Movie Average` on the last row to get the average of each movie in each column. The last column for the last row will be `NA`. Another line named `Movie Avg - Mean Movie` is going to output `Movie Average` subtracted by `globalMean`. After splitting the two tables, I will utilize the `knitr` and `kableExtra` packages to output the tables. The `wideRatings` will be utilized to calculate the mean rating for each critic, and will tkae the difference from `globalMean`. The columns that are for `userAvg` and `globalMeanUserAvg` will be renamed to `Average Rating` and `User Average - Mean Movie` respectively (combines the previous tables in a condensed format).

```{r echo=TRUE}

wideRatings <- movieRatings %>%
  select(userName, movieName, rating) %>%
  pivot_wider(names_from = movieName, values_from = rating)

globalMean <- mean(movieRatings$rating, na.rm = TRUE)

wideRatings <- wideRatings %>%
  rowwise() %>%
  mutate(userAvg = mean(c_across(where(is.numeric)), na.rm = TRUE),
         globalMeanUserAvg = userAvg - globalMean)

wideRatings <- wideRatings %>%
  rename(Critic = userName, 
         `Average Rating` = userAvg, 
         `User Average - Mean Movie` = globalMeanUserAvg)

moviecolumns <- colnames(wideRatings)[2:(ncol(wideRatings) - 2)]
firsttenMovies <- moviecolumns[1:10]
lasttenMovies  <- moviecolumns[11:20]

firstMovieset <- wideRatings %>%
  select(Critic, all_of(firsttenMovies), `Average Rating`, `User Average - Mean Movie`)

secondMovieset <- wideRatings %>%
  select(Critic, all_of(lasttenMovies), `Average Rating`, `User Average - Mean Movie`)

firstMovieSetAvg <- movieRatings %>%
  filter(movieName %in% firsttenMovies) %>%
  group_by(movieName) %>%
  summarise(movie_avg = mean(rating, na.rm = TRUE)) %>%
  pivot_wider(names_from = movieName, values_from = movie_avg) %>%
  mutate(Critic = "Movie Average", `Average Rating` = globalMean, `User Average - Mean Movie` = NA)

secondMovieSetAvg <- movieRatings %>%
  filter(movieName %in% lasttenMovies) %>%
  group_by(movieName) %>%
  summarise(movie_avg = mean(rating, na.rm = TRUE)) %>%
  pivot_wider(names_from = movieName, values_from = movie_avg) %>%
  mutate(Critic = "Movie Average", `Average Rating` = globalMean, `User Average - Mean Movie` = NA)

firstMoviesetGAvg <- firstMovieSetAvg %>%
  mutate(across(all_of(firsttenMovies), ~ . - globalMean)) %>%
  mutate(Critic = "Movie Avg - Mean Movie", `Average Rating` = NA, `User Average - Mean Movie` = NA)

secondMoviesetGAvg <- secondMovieSetAvg %>%
  mutate(across(all_of(lasttenMovies), ~ . - globalMean)) %>%
  mutate(Critic = "Movie Avg - Mean Movie", `Average Rating` = NA, `User Average - Mean Movie` = NA)

firstMovieset <- bind_rows(firstMovieset, firstMovieSetAvg, firstMoviesetGAvg)
secondMovieset <- bind_rows(secondMovieset, secondMovieSetAvg, secondMoviesetGAvg)

```

### First 10 Movies 

This table outputs the first 10 movies with the Global Base Estimate:

```{r echo = TRUE}

firstMovieset %>%
  kable(digits = 2, align = 'lcccccccccccc') %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

```

### Last 10 Movies

This will output the last 10 movies with the Global Base Estimate:

```{r echo = TRUE}

secondMovieset %>%
  kable(digits = 2, align = 'lcccccccccccc') %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

```

## Conclusion

My initial plan was to utilize the large dataset that I had from [MovieLens](https://grouplens.org/datasets/movielens/), but this would have took a bit more time and I am still working with tables to make it better. I did experience a similar issue when I worked with GTFSR with my assignment in undergrad, but as I progress in this course, I believe like this can improve significantly. 